The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
new bibitems, some work on the discussion section

\paragraph{Human-centered perspective on anthropomorphism}

@@

\subsubsection{Initialization Phase}
\label{sec:initialization}

\paragraph{Initial Capital of Anthropomorphism}
\label{sec:ica}

@@

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined application context of the robot, or in which situation the interaction occurs, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in \citet{joosse_what_2013}. The authors showed for instance that when the
same robot (Nao) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different personalities to the robot. In general,
a robot which is imagined in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This was also found in an analysis of anthropomorphic language in online forum discussions about the AIBO, Roomba, and iPad \citep{fink_anthropomorphic_2012}. Regardless of the conversation topic, more anthropomorphic language was found in the AIBO forum. However, this was not only due to the zoomorphic (pet-like) shape of AIBO but also due to its  playful purpose. The idea that the purpose of a robot are crucial factors, also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}. Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities~\citep{baxter2013do}).

\paragraph{The Novelty Effect}
\label{sec:noveltyeffect}

\fxfatal{SL: Julia, do you have smthg to provide an initial account of what the
novelty effect is?}

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households. The observed scenario describes how an elderly
woman (71 years, living on her own with a little dog in a small row house) was
setting up and using the robot for the first time for about 10 minutes.

The woman (E.) is reading aloud parts of the ``quick start manual'': \emph{``When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area.''} She laughs and turns to the robot, talking directly to
it \emph{``Today you can light up darling!''}. E. carries Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{``It has to start from here, this is where
you enter the house!''}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone ``Uh-Oh'' sound followed by a narrated voice that starts off with ``Error
1'']. E. is very surprised \emph{``What, it is even talking to me?''}, lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{``Nice!''} or \emph{``Good job!''} or \emph{``Bravo!''}. Once, when Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{``There you forgot something, what is with over
there? Look, there you still have something to clean up!''}, and she points at
where she wants Roomba to go. Further, she uses deictic gestures and tells the
robot: \emph{``No, go this way! Come over here!''}. The woman keeps observing the
robot and wonders about its intentions, asking it directly: \emph{``Why do you
leave the kitchen? There is still something there on the fringe!''}. E. turns to
the experimenter, and tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{``darling''}
/ \emph{``she''}); praising the robot (\emph{``Good job!''}); using pointing
gestures toward the robot, or try to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{``she loves the hall!''}).

During the following visits at the woman's place, she did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot.

\subsubsection{Familiarization Phase}
\label{sec:familiarization}

\emph{Familiarization}, in this context, means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect at section~\ref{sec:cognition-neuroscience}.

@@

\paragraph{Disruptive behaviors}

@@

As illustrated on Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviors, and their cognitive correlates.

\subsubsection{Stabilization Phase}
\label{sec:stabilization}

@@

\paragraph{Stabilized Level of Anthropomorphism}

@@

@@

This section provide an initial interpretation of anthropomorphism in
terms of \emph{cognitive correlates}: how human cognition can explain the
dynamic nature of our interactions with robots, and what are the main cognitive
stages upon which researchers can build to design better, deeper human-robot
interactions.

The next sections introduces a model of the cognitive stages involved in
human-robot interaction, that can be interpreted in parallel to the model of the
dynamics of anthropomorphism.

We also discuss two specific aspects: the cognitive impact of disruptive
behaviors (section~\ref{sec:disruptive}) on mutual modelling, and how our models
relate to the \emph{In-the-moment} versus \emph{Reflective} perspectives
introduced by~\citet{takayama_perspectives_2012}.

@@

\subsubsection{Cognitive Processes and Stages}

@@

The model of anthropomorphism we propose is still a work in progress. In
particular, its implications regarding the design of (social) robots and
human-robot interaction are still to be refined.

\subsection{Effect of Time and Familiarization}

\fixme{Discuss here the general findings of the paper: the global shape of the
model of the dynamics, ...}

The \textit{context of use} is related to the purpose (and functionality) of the
robot and influences the user interaction experience. With \textit{time}, we
refer to long-term interaction with the robot, which is related to what has been
described as \textit{novelty effect} but also accounts for the user getting used
to the robot. Consequently, we propose that anthropomorphism is not static but
likely to change, due to what we call \textit{dynamics}, in space and time.

As outlined in section \hl{XX}, the tendency to anthropomorphize is motivated by
a person's wish to make sense of an agent which might be difficult to understand
(in its functionality or behavior, for instance). In turn, a user might ascribe
intentions or emotions to the system because the systems output was unexpected
for the user. Based on this explanation, we estimate that when the user has
familiarized herself / himself with the system, thus reached the point of when
the system is usable and explainable, the tendency to anthropomorphize
decreases. Therefore, our second hypothesis is as follows:

\begin{itemize}
    \item \textbf{Hypothesis 1} A person's tendency to anthropomorphize a robot
        is impacted by the context of use and purpose of the robot. A social
        context and purpose of the robot increases anthropomorphism.
        \textit{(context and purpose of the robot)}
	\item \textbf{Hypothesis 2} A person's tendency to anthropomorphize a robot decreases over time and with growing experience the person has with the robot. \textit{(familiarize oneself with the robot)}
\end{itemize}

Given our first hypothesis from above, this might not hold over all contexts;
more concretely, probably not in a social context of usage.

Smart innovative devices, such as personal domestic robots, might demand from a
human user spontaneous usage of an unfamiliar system. Such situations will
require that a user should be able to build up a mental model quickly, which
particularly holds for novice users \cite{schmitz_concepts_2011}.

Eddy \textit{et al.} points out that familiarity increases the tendency to
anthropomorphize \cite{eddy_attribution_1993}.

"As shown in several psychological experiments [13,24] and pointed out by Watt
[25], familiarity may also ease social acceptance and even tend to increase
people's tendency to anthropomorphize [16]." \cite{duffy_anthropomorphism_2003}

\subsection{Limitations}

The model of anthropomorphism we propose is still a work in progress. In
particular, its implications regarding the design of (social) robots and
human-robot interaction are still to be refined.

>>>>

\paragraph{Human-centered perspective on anthropomorphism\\}

@@

\subsubsection{Initialization Phase\\}
\label{sec:initialization}

\paragraph{Initial Capital of Anthropomorphism\\}
\label{sec:ica}

@@

By taking the \textbf{purpose} of a robot into account, we suggest that the real
or imagined application context of the robot, or in which situation the interaction occurs, impacts how far the robot will be attributed human-like
characteristics. We draw on findings such as presented in \citet{joosse_what_2013}. The authors showed for instance that when the
same robot (Nao) is used in a different task context (cleaning task \emph{vs.}
tour guide), users ascribe different personalities to the robot. In general,
a robot which is imagined in a social, entertaining or playful
context leads to a higher \ICA than a robot which is used for a routine or
focused task (security, rescue, etc.). This was also found in an analysis of anthropomorphic language in online forum discussions about the AIBO, Roomba, and iPad \citep{fink_anthropomorphic_2012}. Regardless of the conversation topic, more anthropomorphic language was found in the AIBO forum. However, this was not only due to the zoomorphic (pet-like) shape of AIBO but also due to its  playful purpose. The idea that the purpose of a robot are crucial factors, also receives support from
Goetz \& Kiesler's work that revealed that people prefer a serious robot for
serious tasks and a less serious robot for more playful
tasks~\citep{goetz_cooperation_2002, goetz_matching_2003}.
However, what about artificial agents that seem to have no purpose? \cite{kaplan_free_2000} discusses the role of uselessness in the design of robots. He argues that artificial pets such as the AIBO, have no real purpose, in a sense that they do not provide any kind of service, or, in other words \textit{``they are not doing what you tell them to do''} and \textit{``they might refuse the order of its owner''}. It is this very aspect that might increase people's tendency to anthropomorphize the robot. According to Kaplan, it is in our daily use language, that we tend to attribute intentions to devices that are not doing their job well. For instance, we do not develop any kind of relationship with our computer, washing machine or TV set when they work properly. The author argues that it is only when they start dis-functioning, when they show that they can act differently that what we ordered, that we are ready to give them some kind of ``intentionality''.

 Also, we suggest that
the environmental context in which people experience and interact with the robot
impacts the \ICA. For instance, several friends interacting simultaneously with
the robot might lead to increased \ICA for each of the users, due to increased
human-human social interactions (the robot might be perceived to be part of the
social interaction, and in turn attributed human-like
qualities~\citep{baxter2013do}).

\paragraph{The Novelty Effect\\}
\label{sec:noveltyeffect}

Novelty effects have been described with interactive technologies in general \citep{rogers_diffusion_1995} and also in several long-term HRI studies in workplaces \citep{huttenrauch_fetch-and-carry_2003,mutlu_robots_2008}, in schools and public places \citep{gockley_designing_2005,kanda_interactive_2004,kanda_communication_2005}, or elder care centers \citep{sabelli_conversational_2011}. Further, also with domestic robots it has been found that people's interest, engagement, and fascination with the robot decreased over time \citep{fink_living_2013,fernaeus_how_2010,sung_robots_2009}.

To illustrate how the novelty of a robot can evoke anthropomorphic effects, we
present an example from our long-term ethnographic study with Roomba vacuum
cleaning robots in households \citep{fink_living_2013}. The observed scenario describes how an elderly
woman (71 years, living on her own with a little dog in a small row house) was
setting up and using the robot for the first time during about 10 minutes.

The woman (E.) is reading aloud parts of the ``quick start manual'': \emph{``When
Roomba senses dirt, the blue dirt detect light is lit and Roomba cleans more
intensely in that area.''} She laughs and turns to the robot, talking directly to
it \emph{``Today you can light up darling!''}. E. carries the Roomba to the entrance
door and puts it on the floor [floor type: tiles, partially covered with little
Persian carpets], she explains \emph{``It has to start from here, this is where
you enter the house!''}. After pressing the CLEAN button, the robot quickly gets
stuck with the carpet in the entrance hall, consequently it reports error 1 [
two-tone \textit{``uh-oh''} sound followed by a narrated voice that starts off with \textit{``error
1''}]. E. is very surprised \emph{``What, it is even talking to me?''}, she lifts the
robot off and puts it back for cleaning. At several occasions, while watching
how the robot moves around and cleans, E. makes compliments to the robot:
\emph{``Nice!''} or \emph{``Good job!''} or \emph{``Bravo!''}. Once, when the Roomba turns
right in front of some crumbs on the kitchen floor and moves away without
cleaning them up, E. says: \emph{``There you forgot something, what about over
there? Look, there you still have something to clean up!''}, and she points at
where she wants the Roomba to go. Further, she uses deictic gestures and tells the
robot: \emph{``No, go this way! Come over here!''}. The woman keeps observing the
robot and wonders about its intentions, asking it directly: \emph{``Why do you
leave the kitchen? There is still something there on the fringe!''}. E. turns to
the experimenter, and tells her: \emph{``She [the robot] loves the hall!''}.

Besides this behavior of generally anthropomorphizing the robot, we observe the
following concrete instances as anthropomorphic effects in this scenario:
talking directly to the robot, or asking it questions (though it doesn't
recognize speech); giving it nick-names or ascribing a gender (\emph{``darling''}
/ \emph{``she''}); praising the robot (\emph{``Good job!''}); using pointing
gestures toward the robot, or trying to direct it around with commands; attributing
a personal preference / intention to the robot (\emph{``she loves the hall!''}).

During the following visits at the woman's place, E. did not talk to the robot
anymore, and generally behaved in a less anthropomorphizing way. The observed
effects were likely due to the novelty of the robot, and the fact that E. was not familiar with the robot and its behavior.
As \cite{persson_anthropomorphism_2000} names it: \textit{``Anthropomorphism is a way of making sense of complex behavior in the world around us.''}. Thus, as soon as this complexity is reduced, also anthropomorphism might be reduced.

\subsubsection{Familiarization Phase\\}
\label{sec:familiarization}

In this context, \emph{familiarization} means \emph{to get acquainted with}.
It is not to be confused with the \emph{familiarity thesis} that relates to the
projection of already known cognitive models onto the robot. We discuss this other
aspect at section~\ref{sec:cognition-neuroscience}.

@@

\paragraph{Disruptive behaviors\\}

@@

As illustrated in Figure~\ref{fig:dynamics}, our model represents
\emph{disruptive behaviors} as either local increases or local decreases of
anthropomorphic effects in the familiarization phase (and at a lesser extend, in
the stabilization phase): because such behaviors are unexpected, a human
observer may interpret them either as the result of a richer deliberative
process, which in turn leads to the supposition of complex cognitive skills, or
conversely as a failure, reminding the human that the robot is ``just a
machine''. Section~\ref{sec:disruptive} discusses more in depth the impact of
the user's perception of disruptive behaviors, and their cognitive correlates.

To illustrate how a disruptive behavior of a robot can evoke anthropomorphic effects, we present again an example from our long-term ethnographic study with Roomba vacuum cleaning robots in households \citep{fink_living_2013}. The observed scenario describes how a family (young parents, 26 and 28 years old, with 3 little children, 5 and 4 years, and 6 months old, living in a house with garden on the countryside) were using the robot during the first week.

The father (M.) always lets Roomba clean up the ground floor after the family finishes dinner and they all go upstairs to put the children to bed. When the robot would have finished its cleaning cycle, the parents would later just find it docked at its charging station. It used to be like this every evening. However, one day when the father comes downstairs, Roomba is not at its station. He says: \emph{``Hey, Roomba, where are you? Why did you not want to got to sleep last night?''}. After some searching he found it stuck under the sofa and wondered why this had happened.

What is the observed anthropomorphic effect in this scenario? Usually, the father did not talk to the robot or praise it. However, now that the robot did something unexpected (it was not at its usual place) he started wondering about its ``intention'' and directly asked the robot for its reasons to not \textit{``go to sleep''}.

\subsubsection{Stabilization Phase\\}
\label{sec:stabilization}

@@

\paragraph{Stabilized Level of Anthropomorphism\\}

@@

\fixme{JF: Severin, we should at some point mention what the ``expectation mismatch'' is ... I haven't found that so far ... but I am also not sure about how to explain this. I think that, the we already took into account the user's expectations in the human-factor.}

@@

This section provides an initial interpretation of anthropomorphism in
terms of \emph{cognitive correlates}: how human cognition can explain the
dynamic nature of our interactions with robots, and what are the main cognitive
stages upon which researchers can build to design better, deeper human-robot
interactions.

The next sections introduce a model of the cognitive stages involved in
human-robot interaction, that can be interpreted in parallel to the model of the
dynamics of anthropomorphism.

We also discuss two specific aspects: the cognitive impact of disruptive
behaviors (section~\ref{sec:disruptive}) on mutual modelling, and how our models
relate to the \emph{in-the-moment} versus \emph{reflective} perspectives
introduced by~\citet{takayama_perspectives_2012}.

@@

\subsubsection{Cognitive Processes and Stages\\}

@@

The model of anthropomorphism we propose is still a work in progress. In
particular, its implications regarding the design of (social) robots and
human-robot interaction are still to be refined but we hope this article helps encourage an open dialogue between those who do research of anthropomorphism.

\subsection{Effect of time / familiarization on anthropomorphism}

\fixme{Discuss here the general findings of the paper: the global shape of the
model of the dynamics, ... // JF: yep, I agree, we should also come back to our initial hypotheses}

We have proposed that time, or in other words, the user's familiarization with the non-human agent, models how anthropomorphism evolves over time. Our general assumption is that with continued interaction, and in turn growing user experience, the system becomes more predictable and consequently, the user's need to anthropomorphize the agent decreases. We would like to discuss this suggestion here.

With \textit{time} as a factor, we
referred to long-term interaction with the robot, which is related to what has been
described as \textit{novelty effect} but also accounts for the user getting used
to the robot.

As outlined before, the tendency to anthropomorphize is motivated by
a person's wish to make sense of an agent which might be difficult to understand
(in its functionality or behavior, for instance). In turn, a user might ascribe
intentions or emotions to the system because the system's output was unexpected
for the user. Based on this explanation, we estimate that when the user has
familiarized herself / himself with the system, thus reached the point of when
the system is usable and explainable, the tendency to anthropomorphize
decreases.

However, there exist views that suggest just the opposite, namely, that familiarization with an agent \textit{increases} the tendency to anthropomorphize \citep{eddy_attribution_1993}. Similarly, \cite{duffy_anthropomorphism_2003} reports that several psychological experiments and human-robot interaction experiments showed that familiarity may also ease social acceptance and even tend to \textit{increase} people's tendency to anthropomorphize.

Consequently, we have to accept that our model of the \emph{Dynamics of Anthropomorphism} needs to be verified, to see how far our hypotheses can hold. In particular, the effect of the user's familiarity with the system needs to be studied over a longer period of time.

%\begin{enumerate} 
%\item ~A person's tendency to anthropomorphize a robot
%    is based on several factors, including (but not limited to) (1) the
%    design of the robot, (2) the personal characteristics of the human user,
%    and (3) the context and situation of use. For instance, a social context
%    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
%    as a multi-factor phenomenon}.
%
%\item ~A person's tendency to anthropomorphize a robot
%    evolves in a non-linear way over time and with growing experience that the
%    person has with the robot. For instance, a more experienced user will
%    perceive the robot as more familiar and predictable, and in turn
%    anthropomorphizes the robot less: \textit{anthropomorphism as a dynamic
%    phenomenon}.
%
%\end{enumerate}

%
%Smart innovative devices, such as personal domestic robots, might demand from a
%human user spontaneous usage of an unfamiliar system. Such situations will
%require that a user should be able to build up a mental model quickly, which
%particularly holds for novice users \cite{schmitz_concepts_2011}.
